%===================================================================================================
\section{Output Model: Representing Tests} %++++++++++++++++++++++++++++++++++++++++++++++++++++++++
  \label{Output Model: Representing Tests} %++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================


\improve


\begin{justify}
To make failure feedback consistent across different testing mechanisms, Assertify/Checkify use a
common \emph{output model} that represents the essential components of a test outcome in a
structured form. The model is designed around the minimal information that is directly actionable
for novices: the tested expression and an explicit expected/actual comparison.
This directly operationalises the requirements derived from the problem analysis chapter.
\end{justify}


%===============================================================================
\paragraph{Core fields.}
    \label{Core fields.}
At the centre of the model are three fields:

\begin{enumerate}
    \item \textbf{Tested expression}: a textual representation of what was evaluated
        (e.g., \texttt{digitSum 123N = 6N}), intended to be readable without requiring knowledge of
        the test harness.
    \item \textbf{Expected value}: the reference value implied by the specification or by the
        right-hand side of an equality.
    \item \textbf{Actual value}: the value produced by evaluating the implementation under test.
\end{enumerate}

This triad supports the stable reading strategy targeted by Assertify: identify the
expression, compare expected vs.\ actual, then localise the relevant part of the implementation.


%===============================================================================
\paragraph{Readable vs. technical renderings.}
    \label{Readable vs. technical renderings.}
In practice, expression strings produced by tooling
may include module qualifications or syntactic artifacts that are not helpful for novices.
The model therefore distinguishes between a \emph{full} expression rendering (useful for
traceability in the pipeline) and an optional \emph{simplified} rendering intended for
student-facing output. This design is reflected in the before/after examples already included in
the draft, where both the fully qualified and simplified versions are displayed.


%===============================================================================
\paragraph{Optional context fields.}
    \label{Optional context fields.}
The model also supports optional context fields that may be
included depending on the test mechanism and configuration:

\begin{itemize}
    \item \textbf{History}: a sequence of successfully evaluated expressions leading up to a
        failure, intended to preserve operational context in multi-step tests.
    \item \textbf{Property-based inputs}: original and shrunk counterexamples for property-based
        failures, enabling students to focus on the minimal failing case while retaining access to
        the unshrunk case for additional context.
    \item \textbf{Diagnostic metadata}: a message and stack trace for traceability and for
        tutor/debugging use, while keeping this information secondary to the student-facing summary.
\end{itemize}


%===============================================================================
\paragraph{Serialisation and downstream use.}
    \label{Serialisation and downstream use.}
A structured output model also enables alternative
presentations beyond plain console text. The project notes explicitly mention serialisable outputs
(e.g., JSON) as a basis for visualisation and for integrating reports into external tooling
(such as IDE support). While the thesis primarily discusses the human-readable rendering,
the model is designed to remain usable in such downstream scenarios.


%===================================================================================================
% EOF ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================