%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Problem Analysis: Issues with Existing Error Messages} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sctn:errorMessages} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \todo{}


In this chapter I will explain known issues for novices when working with error messages.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples of Current Error Messages} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \todo{Show examples of unhelpful or ambiguous error messages.}

Error messages often require the programmer to not only know what was actually tested but how it was tested as well.
For instance are characteristics of data structures sometimes tested multiple times between altering the stored data (e.g. adding or removing elements).
To then comprehend the error message of such a test fully would require one to know every alteration done before the test/assertion failed to reproduce and retrace the steps
and by that find the flaw in their implementation.\\[5pt]
Furthermore are some error messages formatted in a manner such that essential data for novices is broken up and/or hidden.
In particular does \href{https://learn.microsoft.com/en-us/dotnet/api/microsoft.visualstudio.testtools.unittesting.assert.areequal?view=mstest-net-3.8}{\texttt{Assert.AreEqual}}
not show which method/expression was tested when the error occurred and rather focues on the input.
Without including the name of the tested method in the signature of the test method a student would not be able to connect the test results with any particular part of their code which renders the error message basically useless.
Even if the name is included you still have to connect the separate pieces to reconstruct the explicit method call that caused this error message.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\input{matter/error-messages/digitSum.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\input{matter/error-messages/sortedDigits.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Information Gaps in Error Outputs} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \todo{Analyze what critical information is missing in these messages.}


One thing error messages lack is information about the context of a failed test.
Consider an implementation of a singly linked list with no duplicates. The linked list contains the attributes: pointer to both the first and last node of the list and the size of list.
Each node contains a value and a pointer to the next element.

\begin{tcbtikz}{Illustration singly linked list}{singlyLinkedList}{}
    % \draw[step=.5cm,gray,very thin] (-6, -6) grid (6, 6);
    % \path[draw=black, fill=black] (0, 0) circle (.05cm);
    \begin{scope}[xshift=-5mm]
        \node[font=\large] at (0, 1) {\textbf{Data:} \(\{\, 1,2,\dotsc,10 \,\}\)};
        \node[state, fill=white] at (-5, 0) (n1) {1};
        \path[-{Stealth[length=2mm,width=2mm]}, red] (-5, -1) edge node [below, yshift=-2mm] {\textbf{first}} (n1);
        \node[state, fill=white] at (-3, 0) (n2) {2};
        \draw[-{Stealth[length=2mm,width=2mm]}] (n1) edge node [above,yshift=1mm] {next} (n2);
        \draw[-{Stealth[length=2mm,width=2mm]}] (n2.east) to node [above,yshift=1mm] {next} ([xshift=1cm]n2.east);
        \path[draw=black, fill=black] (-1, 0) circle (.05cm);
        \path[draw=black, fill=black] (-.5, 0) circle (.05cm);
        \path[draw=black, fill=black] (0, 0) circle (.05cm);
        \node[state, fill=white] at (2, 0) (n9) {9};
        \draw[-{Stealth[length=2mm,width=2mm]}] (.5, 0) to node [above,yshift=1mm] {next} (1.5, 0);
        \node[state, fill=white] at (4, 0) (n10) {10};
        \draw[-{Stealth[length=2mm,width=2mm]}] (n9) edge node [above,yshift=1mm] {next} (n10);
        \draw[-{Stealth[length=2mm,width=2mm]}] (n10.east) to node [above,yshift=1mm] {next} ([xshift=1cm]n10.east);
        \path[-{Stealth[length=2mm,width=2mm]}, red] (4, -1) edge node [below, yshift=-2mm] {\textbf{last}} (n10);
        \node[font=\large] at (6, 0) {null};
        \node[font=\large, red] at (0, -1) {\textbf{size} = \(10\)};
    \end{scope}
\end{tcbtikz}

\noindent Furthermore consider the implemented methods:
\begin{itemize}[nosep]
    \item \fsharp|pushFront: 'a -> LinkedList -> unit|
    \item \fsharp|pushBack: 'a -> LinkedList -> unit|
    \item \fsharp|insertAtIndex: 'a -> int -> LinkedList -> unit|
    \item \fsharp|remove: 'a -> LinkedList -> unit|
\end{itemize}

\noindent These methods would need to be tested for an empty and non-empty list.
As well as whether combinations of method application do not break the list, namely: Inserting and element and then removing it, removing elements until the list is empty and so on.
% For this illustration I will more or less formulate pseudo-code.

\todo{Illustrate how tests for this example might leave gaps because it misses the information of what was actually tested.
How many elements were inserting before the removal of an element somehow failed?
Introduce idea of a \textbf{history}, so that every error message comes with a history of the executed operations before something failed.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Challenges for Novices in Deriving Test Information} %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\todo{Explain why deciphering test expectations is hard for beginners. Elaborate that the information is spread across the entire error message and one has to derive essential information.}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EOF %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%