%===================================================================================================
\chapter{Introduction} %++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
  \label{Introduction} %++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================


\improve


\begin{justify}
All programmers have to start at the beginning: What is programming? What is required to write and execute code? What are variables and methods and how are they used?
From there, learners gradually encounter key concepts and algorithms, and later data structures such as maps and sets. However, moving from basic syntax to
independently writing correct programs is a long process, which also includes implementing corresponding unit tests to validate solutions.

\skipM Understanding and interpreting error messages is a fundamental challenge for novice programmers. Error messages serve as an essential form of feedback,
enabling programmers to diagnose issues in their code. In practice, standard diagnostics - especially those produced by compilers and testing frameworks -
are often technical, verbose, and difficult for beginners to interpret.

\skipM This becomes particularly apparent in introductory programming courses where students receive exercise templates with predefined tests and
are expected to implement missing functions or data structures. In such settings, learners are frequently unfamiliar with unit testing and
are immediately confronted with stack traces and framework-specific output formats.

\skipM In \fs, multiple testing frameworks and assertion libraries are available. A commonly used option is
\href{https://learn.microsoft.com/en-us/dotnet/api/microsoft.visualstudio.testtools.unittesting}{Microsoft.VisualStudio.TestTools.UnitTesting}. For example, using
\href{https://learn.microsoft.com/en-us/dotnet/api/microsoft.visualstudio.testtools.unittesting.assert.areequal?view=mstest-net-3.8}{\texttt{Assert.AreEqual}}
can yield the following output on failure:
\begin{verbatimpanel}[error message, hl={2-3}, hlcolor=red!90]{Example Ouput of \texttt{Assert.AreEqual}}
Assert.AreEqual failed. Expected:<[1; 2; 4]>. Actual:<[1; 4; 2]>.
    at MapTests.TestsMapSortedList.Tests.testSetBeispiele() in /code/TestsMapSortedList.fs:line 53
    at System.RuntimeMethodHandle.InvokeMethod(Object target, Void** arguments, Signature sig, Boolean isConstructor)
    at System.Reflection.MethodBaseInvoker.InvokeWithNoArgs(Object obj, BindingFlags invokeAttr)
\end{verbatimpanel}

\noindent For more advanced programmers, this output is usually sufficient: they can quickly identify the relevant mismatch and ignore the remaining details.
See Section \hl{TODO} for a discussion of common issues in diagnostic messages and strategies for extracting key information.

\skipM In my experience as a tutor for a first-semester lecture, I frequently observe students struggling with such messages, unsure what went wrong and how to fix their code.
Instead of guiding them toward a solution, poorly structured error messages can lead to frustration and trial-and-error debugging,
which may hinder learning and delay progress on key programming concepts.

\skipM Didactically, complex concepts are typically introduced gradually, often by reducing incidental complexity at the beginning:

\skipS\begin{itemize}[nosep]
    \item Lightweight IDE's are often preferred to heavy, feature-rich IDE's
    \item Simplistic I/O is introduced before full input validation only valid inputs and error handling (e.g., malformed inputs, exceptions, and missing values).
\end{itemize}

\skipS The same  Diagnostic verbosity can be staged for testing: Initially showing only the most relevant information and revealing additional technical details as students gain experience.
This thesis explores how a similar scaffolding approach can be applied to testing feedback. By providing more minimalistic and structured output,
the goal is to help students interpret failures more effectively and support a more efficient and engaging learning process.
\end{justify}


%===============================================================================
% PROBLEM STATEMENT %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===============================================================================


\input{matter/main/introduction/sections/problem-statement.tex}


%===============================================================================
% RESEARCH QUESTIONS %++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===============================================================================


\input{matter/main/introduction/sections/research-questions.tex}


%===============================================================================
% THESIS STRUCTURE %++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===============================================================================


\input{matter/main/introduction/sections/thesis-structure.tex}


%===================================================================================================
% EOF ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================