%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Assertify: A Library for Improving Error Messages} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


Assertify describes a library of various methods and classes to improve error messages during testing.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Objectives and Design of Assertify} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \todo{State the goals and design principles behind Assertify.}

\noindent Assertify incorporates functionalities to enhance the debugging and testing experience by providing clearer, more informative error messages.
Traditional assertion failures often yield cryptic outputs, making it difficult to identify the root cause of an issue.
Assertify addresses this by offering detailed, structured feedback, including expression simplifications, expected vs. actual values, and reduction history.
The design principles behind Assertify focus on:
\begin{enumerate}[label=\textbf{(\arabic*)}]
    \item \textbf{Clarity} -- Displaying test failures in a structured and easily readable format.
    \item \textbf{Context} -- Providing additional context through expression reductions and history tracking.
    \item \textbf{Usability} -- Ensuring that error messages are helpful for students and developers alike.
    \item \textbf{Flexibility} -- Supporting different levels of verbosity for various debugging needs.
\end{enumerate}
By integrating these principles, Assertify streamlines the debugging process and improves the efficiency of error resolution.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation Details and Wrappers} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \todo{Provide technical details on how Assertify wraps tests and improves output.}

Assertify is implemented as a wrapper around standard test assertions, enriching the output with structured details. The core mechanism involves capturing the tested expressions, simplifying them where possible, and formatting the results into a well-structured report.
Key implementation details include:
\begin{itemize}
    \item \textbf{Expression Reduction:} Assertify simplifies complex expressions, stripping unnecessary details while preserving meaningful structure.
    \item \textbf{History Tracking:} The system logs successfully evaluated expressions leading up to a failure, providing insight into the progression of computations.
        This is especially helpful if a succession of operations are executed on the same data structure and said data structure is correctly implemented for the majority of operations
        and only fails when a specific edge case occurs (e.g. an (almost) full linked list encounters a out-of-bounds index error).
        Then the history of operations might be helpful for novices to recreate the error because the order of operations is crutial to whether or not a operation fails.
    \item \textbf{Verbose and Compact Modes:} Users can toggle between detailed explanations and minimalistic outputs.
    \item \textbf{Enhanced Formatting:} Utilizing structured markers and color-coded elements, Assertify highlights critical differences between expected and actual results.
\end{itemize}
By wrapping around existing test frameworks, Assertify seamlessly integrates into existing workflows, enhancing error visibility without requiring significant changes to test code.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Error Message Improvements (Before and After Comparisons)} %%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \todo{Compare previous error messages with Assertify's outputs.}


Traditional assertion failures often return extremely detailed messages that require additional effort to decipher.
Assertify enhances this by providing detailed comparisons. 
The enhanced output not only shows expected vs. actual values but also presents the exact expression tested, making debugging significantly easier.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 1 --- \texttt{digitSum}} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\todo{}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Example Tests --- \texttt{digitSum}} %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{todolist}
    \item Compare old and new test results (old: \ref{fsharp:{outputExampleTestsDigitSum}})
\end{todolist}


\begin{fsharpoutput}%
{Assertify output of examples tests for \texttt{digitSum}}%
{assertifyOutputExampleTestsDigitSum}
-------------------- TESTED EXPRESSION -----------------------
Tested/Failed expression:   StudentSubmission.digitSum 123N
|{\color{blue}Simplified expression:      digitSum 123N}|
------------------- EXPECTED AND ACTUAL ----------------------
|{\color{green!70!black}Expected result:            6N}|
|{\color{red}Actual result:              7N}|
\end{fsharpoutput}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Random Tests --- \texttt{digitSum}} %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{todolist}
    \item Compare old and new test results (old: \ref{fsharp:{outputRandomTestsDigitSum}})
\end{todolist}


\begin{fsharpoutput}%
{Assertify output of examples tests for \texttt{digitSum}}%
{assertifyOutputExampleTestsDigitSum}
-------------------- TESTED EXPRESSION -----------------------
Tested/Failed expression:   StudentSubmission.digitSum n
|{\color{blue}Simplified expression:      digitSum 0N}|
------------------- EXPECTED AND ACTUAL ----------------------
|{\color{green!70!black}Expected result:            0N}|
|{\color{red}Actual result:              1N}|
\end{fsharpoutput}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 2 --- \texttt{sortedDigits}} %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\todo{}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Example Tests --- \texttt{sortedDigits}} %%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{todolist}
    \item Compare old and new test results (old: \ref{fsharp:{outputExampleTestsSortedDigits}})
\end{todolist}


\begin{fsharpoutput}%
{Assertify output of examples tests for \texttt{sortedDigits}}%
{assertifyOutputExampleTestsSortedDigits}
-------------------- TESTED EXPRESSION -------------------------
Tested/Failed expression:   StudentSubmission.sortedDigits 1111N
|{\color{blue}Simplified expression:      sortedDigits 1111N}|
------------------- EXPECTED AND ACTUAL ------------------------
|{\color{green!70!black}Expected result:            true}|
|{\color{red}Actual result:              false}|
\end{fsharpoutput}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Random Tests --- \texttt{sortedDigits}} %%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{todolist}
    \item Compare old and new test results (old: \ref{fsharp:{outputRandomTestsSortedDigits}})
\end{todolist}


\begin{fsharpoutput}%
{Assertify output of examples tests for \texttt{sortedDigits}}%
{assertifyOutputExampleTestsSortedDigits}
-------------------- TESTED EXPRESSION -----------------------
Tested/Failed expression:   StudentSubmission.sortedDigits n
|{\color{blue}Simplified expression:      sortedDigits 33N}|
------------------- EXPECTED AND ACTUAL ----------------------
|{\color{green!70!black}Expected result:            true}|
|{\color{red}Actual result:              false}|
\end{fsharpoutput}
    

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usability and Design Considerations} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% \todo{Discuss design decisions to enhance clarity and usability for students.}

Assertify is designed with a strong focus on usability, particularly for students learning programming and debugging. Key considerations include:
\begin{itemize}
    \item \textbf{Readable Error Reports:} Formatting failures in a structured, natural-language format ensures accessibility.
    \item \textbf{Minimal Intrusion:} Assertify seamlessly integrates with existing test frameworks without requiring major changes.
    \item \textbf{Customizability:} Users can adjust verbosity levels and enable/disable history tracking based on their needs.
    \item \textbf{Color Coding and Annotations:} Highlighting key differences improves readability and makes debugging more intuitive.
\end{itemize}
These design choices collectively improve the debugging experience, reducing frustration and helping developers quickly understand and fix issues.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EOF %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%