%===================================================================================================
\section{Context and Assumptions} %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
  \label{Context and Assumptions} %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================


\improve


\begin{justify}
The analysis in this chapter is situated in introductory programming exercises where students
implement functions according to a specification and receive feedback through compilation and
automated tests. Students typically iterate by running tests locally or by submitting to an
assessment system then inspecting the returned compiler and test feedback. A key assumption is that
students primarily control the \emph{solution code}, while the tests may be partially or fully
hidden. Consequently, the error message becomes the primary artifact from which learners must derive
actionable debugging information.

The examples in this thesis focus on the \fs ecosystem and a test setup that combines (i)
example-based tests using Unquote and (ii) unit and property-based tests executed via the MSTest
framework and FsCheck. The analysis assumes that novice students have limited familiarity with
testing conventions such as stack traces, reflection-based invocation, and framework-specific
formatting. They may also struggle to distinguish between information that is actionable for
correcting their own logic and information that is mainly relevant for diagnosing the test
infrastructure.

Finally, the chapter assumes that effective feedback should reduce reconstruction effort.
Barik et al.\ show that developers spend substantial time attending to error message regions and
that difficulty in reading error messages predicts task performance \cite{barik}. While their
participants are not restricted to novices, the findings motivate presenting relevant failure
information in a structured, easy-to-scan form.
\end{justify}


\begin{comment}
	This thesis is situated in an introductory programming setting in which students solve small-to-medium programming exercises and receive automated feedback from compilation and testing.
	The tasks are designed to be assessable automatically, and correctness is primarily determined through a combination of example-based unit tests (covering typical and edge cases) and,
	where applicable, property-based tests (covering broader input spaces through generated test cases).

	\skipM The problem analysis adopts the following context assumptions:

	\skipS\begin{itemize}
		\item \textbf{Novice audience:} The primary target group consists of first-semester students who have limited experience with debugging and limited familiarity with framework internals
			(e.g., stack traces, reflection-based test runners, or property-based test shrinking).
		\item \textbf{Feedback pipeline:} Students typically encounter feedback in two stages:
			(i) compilation diagnostics when code does not type-check or parse, and
			(ii) test failure output once code compiles and tests execute.
		\item \textbf{Tooling baseline:} Tests are executed using established infrastructures (unit testing and property-based testing libraries).
			The analysis treats these infrastructures as the baseline whose output is to be \emph{interpreted by students}, not as systems to be redesigned from scratch.
		\item \textbf{Non-goals of this chapter:} This chapter does not attempt to redesign compiler diagnostics.
			Instead, it focuses on the \emph{reporting problem} for test failures, while acknowledging that compiler errors remain a major part of the overall novice experience.
	\end{itemize}

	\skipS In addition, the analysis distinguishes between two kinds of information that appear in typical outputs:

	\skipS\begin{itemize}
		\item \textbf{Essential failure information} needed to understand \emph{what} failed (e.g., the failing expression or behaviour, and the expected vs.\ actual outcome).
		\item \textbf{Auxiliary technical information} that can be useful for advanced debugging but often increases cognitive load for novices
			(e.g., full stack traces, internal framework calls, runner configuration details, reflection invocation frames).
	\end{itemize}

	\skipS\textbf{Checkpoint (to fill with concrete course/tool details):}

	\skipS\begin{itemize}
		\item TODO: Specify the course/exercise setting (language version, IDE/editor setup, where students see output: console, CI/autograder log, IDE test explorer, etc.).
		\item TODO: Summarise the types of tests used (example tests vs.\ random/property tests), including how failures are surfaced to students.
		\item TODO: Clarify what students are allowed to modify (submission file only vs.\ tests visible/hidden), as this affects what output is actionable.
	\end{itemize}
\end{comment}


%===============================================================================
% TODO +++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===============================================================================


% TODO


%===================================================================================================
% EOF ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================