%===================================================================================================
\section{Derived Requirements and Design Implications} %++++++++++++++++++++++++++++++++++++++++++++
  \label{Derived Requirements and Design Implications} %++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================


\improve


\begin{justify}
Based on the issues above, this section derives requirements for novice-oriented test failure
reporting. Each requirement is phrased as a constraint on the \emph{user-facing report}, independent
of any specific implementation technique. The subsequent design chapter then maps these requirements
onto concrete mechanisms (e.g., expression capture, value extraction, and framework wrappers).
\end{justify}


%===============================================================================
\subsection{Requirements} %+++++++++++++++++++++++++++++++++++++++++++++++++++++
     \label{Requirements} %+++++++++++++++++++++++++++++++++++++++++++++++++++++
%===============================================================================


\begin{description}
	\item[\textbf{R1: Stable summary structure.}]
		\begin{justify}
			Every failing test should provide a compact, consistently ordered summary containing at
            least: (i) the tested expression (or an equivalent description of the behaviour under
            test), (ii) the expected value/condition, and (iii) the actual value/observed result.
		\end{justify}
	\item[\textbf{R2: Explicit tested expression.}]
		\begin{justify}
			The report must explicitly identify \emph{what was executed} (e.g., a reconstructed
            function call) rather than relying on the test name or on implicit context.
		\end{justify}

	\item[\textbf{R3: Expected vs.\ actual as the primary comparison.}]
		\begin{justify}
			The expected and actual outcomes must be presented side-by-side (or in a tightly coupled
            pair) in a way that supports rapid visual comparison.
		\end{justify}

	\item[\textbf{R4: Separation of essential and auxiliary information.}]
		\begin{justify}
			Framework internals (stack traces, runner frames, reflection calls) must not obscure the
            essential summary. If included at all, such details should be clearly separated and
            optionally suppressible.
		\end{justify}

	\item[\textbf{R5: Applicability across test paradigms.}]
		\begin{justify}
			The same summary structure should apply to both example-based unit tests and
            property-based tests, even when additional information (e.g., counterexamples) is
            present.
		\end{justify}

	\item[\textbf{R6: Structured representation of property-based counterexamples.}]
		\begin{justify}
			If a property fails, the report must clearly indicate (i) the failing input, (ii) any
            shrunk/minimised counterexample, and (iii) how this input relates to the tested
            expression and observed mismatch.
		\end{justify}

	\item[\textbf{R7: Support for contextual history when needed.}]
		\begin{justify}
			For multi-step/stateful scenarios, the reporting layer should be able to attach an
            execution history (e.g., a sequence of operations) to make the failure reproducible and
            interpretable.
		\end{justify}
\end{description}


%===============================================================================
\subsection{Design implications for Assertify/Checkify} %+++++++++++++++++++++++
     \label{Design implications for Assertify/Checkify} %+++++++++++++++++++++++
%===============================================================================


\begin{justify}
The requirements above imply that a reporting layer must (a) gain access to the tested expression at
runtime, (b) evaluate it to obtain the actual value, (c) retain or compute the expected value, and
(d) serialise both into a stable output model that is independent of any particular baseline
framework.

\skipM In the scope of this thesis, these implications motivate:

\skipS\begin{itemize}
    \item \textbf{Expression capture:} obtain a representation of the tested expression that can be
        displayed to the user (and simplified where appropriate).
    \item \textbf{Value extraction:} evaluate expressions in a controlled way to compute actual
        values and to render values consistently.
    \item \textbf{Wrapper-based integration:} intercept failures from existing unit testing and
        property-based testing frameworks and replace or augment their output without requiring
        changes to the surrounding tooling.
    \item \textbf{Unified output model:} define a single report representation that can be produced
        for different assertion forms and test paradigms.
\end{itemize}

\skipS\textbf{Checkpoint (traceability to later chapters):}

\begin{marker}[med]
    \item TODO: Add a small mapping table from issues $\to$ requirements $\to$ implementation
        mechanisms (one row per requirement).
    \item TODO: Decide which requirements are \emph{must-have} vs.\ \emph{nice-to-have} for the
        first version of Assertify.
    \item TODO: State explicitly which requirements are out of scope (e.g., improving compiler
        diagnostics beyond test reporting).
\end{marker}
\end{justify}


%===============================================================================
\subsection{Completion Checklist for This Chapter} %++++++++++++++++++++++++++++
     \label{Completion Checklist for This Chapter} %++++++++++++++++++++++++++++
%===============================================================================


\begin{marker}[med]
	\begin{itemize}
		\item Add at least one fully worked example for each failure category (compiler error, unit
            test failure, property-based failure).
		\item Add an explicit ``signal vs.\ noise'' annotation for each example output.
		\item Ensure every derived requirement is supported by at least one observed issue
            (traceability).
		\item Add short, properly cited connections to prior work on diagnostic usability and novice
            comprehension (do not overclaim; separate observation from generalisation).
		\item Verify terminology consistency with later chapters (e.g., always use
            \emph{tested expression / expected / actual} in the same order).
	\end{itemize}
\end{marker}


%===============================================================================
% TODO +++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===============================================================================


% TODO


%===================================================================================================
% EOF ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%===================================================================================================