\section*{A Sketch of Testify as a Deeply Embedded DSL}

This note gives a compact, design-oriented sketch of the testing DSL exposed by \texttt{Testify}.
It is intentionally not a full formal semantics and not a parser grammar for arbitrary F\# programs.
Instead, it abstracts the public API into a few core syntactic categories that explain why the DSL is
\emph{deeply embedded}: expectations are explicit F\# values, they can be composed inside the host
language, and operators provide a readable concrete syntax for applying those embedded terms.

\paragraph{Assert DSL.}
The Assert side is expression-left: a quoted expression is validated against one reusable
\texttt{AssertExpectation}.

\[
\begin{array}{rcl}
AExpect & ::= & \textsf{EqualTo}(v)
            \mid \textsf{NotEqualTo}(v)
            \mid \textsf{LessThan}(v)
            \mid \textsf{GreaterThan}(v)
            \mid \textsf{Satisfy}(d, p)
            \mid \textsf{Throws}
            \mid \textsf{DoesNotThrow}
            \mid \textsf{IsTrue}
            \mid \textsf{IsFalse}
            \mid \textsf{IsSome}
            \mid \textsf{IsNone}
            \mid \textsf{Not}(AExpect)
            \mid AExpect \;\textsf{OrElse}\; AExpect
            \mid AExpect \;\textsf{AndAlso}\; AExpect
            \mid \textsf{Any}(AExpect^\ast)
            \mid \textsf{All}(AExpect^\ast)
\end{array}
\]

\[
\begin{array}{rcl}
AExp & ::= & \langle e \rangle \; \mid>? \; AExpect \\
     & \mid & \langle e \rangle \; >>? \; AExpect \\
     & \mid & \langle e \rangle \; =? \; v
            \mid \langle e \rangle \; <>? \; v \\
     & \mid & \langle e \rangle \; <? \; v
            \mid \langle e \rangle \; <=? \; v
            \mid \langle e \rangle \; >? \; v
            \mid \langle e \rangle \; >=? \; v \\
     & \mid & \langle e \rangle \; ^?
            \mid \langle e \rangle \; ^!?
            \mid \langle e \rangle \; ?
            \mid \langle e \rangle \; !? \\
     & \mid & \langle e \rangle \; ||? \; AExpect^\ast
            \mid \langle e \rangle \; \&\&? \; AExpect^\ast
\end{array}
\]

\paragraph{Check DSL.}
The Check side describes relations between a quoted implementation and a reference implementation over
generated inputs. The deeply embedded core is the reusable \texttt{CheckExpectation} value.

\[
\begin{array}{rcl}
CExpect & ::= & \textsf{EqualToReference}
            \mid \textsf{EqualTo}(v)
            \mid \textsf{EqualToReferenceBy}(proj)
            \mid \textsf{EqualToReferenceWith}(cmp)
            \mid \textsf{ThrowsSameExceptionType}
            \mid \textsf{SatisfyWith}(d, p)
            \mid \textsf{SatisfyObservedWith}(d, p)
            \mid CExpect \;\textsf{OrElse}\; CExpect
            \mid CExpect \;\textsf{AndAlso}\; CExpect
\end{array}
\]

\[
\begin{array}{rcl}
CExp & ::= & \langle f \rangle \; |=> \; r \\
     & \mid & \langle f \rangle \; |=>> \; r \\
     & \mid & \langle f \rangle \; |=>? \; (cfg, r) \\
     & \mid & \langle f \rangle \; |=>?? \; (arb, r) \\
     & \mid & \langle f \rangle \; |=>??? \; (CExpect, r) \\
     & \mid & \langle f \rangle \; |=>>? \; (CExpect, r) \\
     & \mid & \langle f \rangle \; ||=>? \; (cfg?, arb?, CExpect?, r)
\end{array}
\]

\paragraph{Declaration layer.}
At the test-hosting layer, Testify wraps MSTest declarations with attributes that trigger Testify's
reporting and result persistence.

\[
\begin{array}{rcl}
TestDecl & ::= & [\texttt{<TestifyClass>}] \; type \; T = \{ MethodDecl^\ast \} \\
MethodDecl & ::= & [\texttt{<TestifyMethod>}] \; member \; m = Body
\end{array}
\]

\paragraph{Mapping to the implementation.}
This sketch corresponds to the current public API as follows.

\begin{itemize}
\item \texttt{AExpect} is represented by the record type \texttt{AssertExpectation<'T>}, whose fields
      describe verification, formatting, explanations, and nested failure details.
\item \texttt{CExpect} is represented by the record type
      \texttt{CheckExpectation<'Args, 'Actual, 'Expected>}, which describes a relation between tested
      and reference behavior for one generated case.
\item The combinators \textsf{OrElse} and \textsf{AndAlso} correspond to
      \texttt{orElse}/\texttt{andAlso} and are surfaced concretely by the operators
      \texttt{<|>} and \texttt{<&>}.
\item \texttt{|>?}, \texttt{|=>}, \texttt{|=>?}, and \texttt{||=>?} belong to the concrete syntax
      layer: they apply embedded expectation values to quoted expressions or quoted implementations.
\item The DSL is therefore ``deep'' mainly at the expectation level; the operators are a readable
      facade over those embedded terms.
\end{itemize}

\paragraph{Interpretation note.}
The grammar above should be read as a compact conceptual model of Testify's testing language. It
captures the core families of terms and applications, but it intentionally omits many specialized
builders such as string-, sequence-, and diff-specific expectations in order to keep the sketch small
and thesis-friendly.
